---
title: "Assignment 2"
author: "Michael Culshaw-Maurer"
date: "1/17/2017"
output: html_document
---

Write up the Ricker model function:
```{r}
RickRoll <- function(nt, r, k){
  nt1 <- nt*exp(r*(1-nt/k))
  return(nt1)
}
```

Using the suggested parameter values:
```{r}
k <- 100 # carrying capacity
n0 <- 50 # starting population size
tf <- 40 # number of time steps to iterate over
```

Now creating the stuff for the for loop simulation:
```{r}
vr <- c(1.5,2.3,2.6,3.0) # creating a vector of r values to use
Nvr <- length(vr) # this is the length of the r value vector
r <- matrix(NA,nrow=tf,ncol=Nvr) # creating an empty matrix with length of the time series,
                                 # and as many columns as r values we'll use
str(r) # checking the structure of the matrix
r[1,] <- rep(n0,Nvr) # putting in our starting value as the first time step entry
for(t in 1:(tf-1)){ # loop through all of the time steps
  r[t+1,] <- RickRoll(r[t,],r=vr,k=k) # use the vector of r values
}

```

Plot our results:
```{r}
# plot over the time series, the results of the simulation, and color them by r value
# also changing the ylim to make a little more room for the legend
matplot(1:tf,r,type = "l",xlab="Time",ylab="Pop Size", ylim=c(0,max(r))+50,col = c("black","blue","red","green"),lty = rep(1.4))
# add a legend with the r value lines marked out
legend("topright", title = "r =", col=c("black","blue","red","green"),legend=c("1.5","2.3","2.6","3.0"),lty = rep(1,4),cex=0.75)
```

Alright, time for a bifurcation diagram.

First, let's clear the values and make some new parameters:

```{r}
rm(r)
rm(t)

k <- 100 # carrying capacity
n0 <- 50 # starting pop size
tf <- 500 # number of time steps to take
```

Creating stuff for the for loop simulation:

```{r}
vr <- seq(from=1.5,to=3.6,by=0.001) # creating a vector of r values to use
Nvr <- length(vr) # this is the length of the r value vector
r <- matrix(NA,nrow=tf,ncol=Nvr) # creating an empty matrix with length of the time series,
                                 # and as many columns as r values we'll use
str(r) # checking the structure of the matrix
r[1,] <- rep(n0,Nvr) # putting in our starting value as the first time step entry
for(t in 1:(tf-1)){ # loop through all of the time steps
  r[t+1,] <- RickRoll(r[t,],r=vr,k=k) # use the vector of r values
}
```

Time to plot:
```{r}
r2 <- r[(tf-39):tf,] # create a new matrix with only the last 40 time points
str(r2) # check the structure to make sure it worked
plot(vr,r2[1,], pch=".", xlab="r", ylab="pop size") # plot the vector of r values against the first
# set of population sizes and time points
for(i in 2:40) points(vr,r2[i,],pch=".") # use a for loop to plot the rest of the time points
```

Aaaaaaand just for fun, let's make a bifurcation diagram with the Lyapunov Exponents plotted on top of it:
```{r}

f=function(r,x)x*exp(r*(1-x)) # function for Ricker
fp=function(r,x)exp(r*(1-x))*(1-r*x) # function for derivative of Ricker
Tf=1000 # how long to calculate lyapunov for

lyap=numeric(Nvr) # vector to store the lyapunov exponents, length of vr used above
x=n0 # initial densities same as for the bifurcation diagram

for(t in 1:Tf){
  x=f(vr,x) # updating the x values for all the r values
  lyap=lyap+log(abs(fp(vr,x))) # summing up all the logs of absolute values of derivatives
}

lyap=lyap/Tf # we want average stretch over time, so we divide by length of the run
```

Time to plot:
```{r}
par(mar=c(5,3,4,3)) # change margins to make room for two y axes
plot(vr,r2[1,], pch=".", xlab="r",ylab=NA) # plot initial bifurcation diagram points
mtext(side=2,line=2,"Pop Size") # add axis label
for(j in 2:40) points(vr,r2[j,],pch=".") # plot the rest of the bifurcation diagram points
par(new=T) # add a new plot on top of the old one
plot(vr,lyap,type="l",col="red",axes=F,xlab=NA,ylab=NA) # plot lyapunov exponents against vr
axis(side=4) # add a second y axis on the right side
mtext(side=4,line=2,"Lyapunov Exponent") # add a second y axis label
abline(h=0,lty=2,col="red") # add a dashed line showing a lyapunov exponent of 0
abline(v=c(2.0,3.1575),lty=2,col=rep("black",2)) # add a few dashed lines to see how bifurcation matches up with lyapunov

```

