---
title: "ECL 233 Assignment 1"
author: "Michael Culshaw-Maurer"
date: "1/11/2017"
output: html_document
---

RickRoll is function to iterate Ricker model one time step. nt is n(t), r is growth rate, k is carrying capacity. Function returns n1, which is n(t+1)

```{r}
RickRoll <- function(nt, r, k){
  n1 <- nt*exp(r*(1-nt/k)) # Ricker equation
  return(n1) # return the next value after N(t)
}
```

BevMo is function to iterate Beverton-Holt model one time step. nt is n(t), r is growth rate, k is carrying capacity. Function returns n1, which is n(t+1)

```{r}
BevMo <- function(nt, r, k){
  n1 <- (exp(r)*nt)/(1+exp(r)*nt/k) # Beverton-Holt equation
  return(n1) # return the next value after N(t)
}
```


saving some baseline values

```{r}
nt <- 1:200 # range of N(t) values (going above the carrying capacity)
r <- 1.2 # growth rate
k <- 100 # carrying capacity
```


generating and saving results for N(t+1) for each model given the same inputs

```{r}
rick <- RickRoll(nt,r,k)
bev <- BevMo(nt,r,k)
```


plotting the results from each model against N(t), adding some color and a legend. I also added a line with slope=1, which helps to show whether N(t+1) is greater or less than N(t).

```{r}
plot(nt, rick, type="l", col="red", xlab="N(t)", ylab="N(t+1)") # generate plot and Ricker curve
lines(nt, bev, col="blue") # add the Beverton-Holt curve
lines(nt,nt, col="black") # add a line with slope=1
legend("bottomright", legend=c("Ricker", "Beverton-Holt", "slope of 1"), col=c("red","blue", "black"), lty = c(1,1,1)) # add a legend for all three lines
```
