---
title: "ECL 233 Assignment 3"
author: "Michael Culshaw-Maurer"
date: "1/26/2017"
output: html_document
---

Read in the teasel data:
```{r}
TS <- as.matrix(read.table("/Users/MJ/GitHub/ECL_233/teasel_stage.txt"))
```


Define a function for creating the projection matrix with values of a and f. I modified mine to include an input for f.
```{r}
A2 <- function(N,f){
  A.temp=TS;
  A.temp[1,7]=f/(1+a*sum(N));
  return(A.temp)
}
```


Time to figure out eigenvalues for the linearization of the model.
```{r}
a <- 0 # we're doing the linearization, so this gets rid of density dependence
f <- 0:75
eigenvalues <- rep(NA,length(f)) # make an eigenvalues vector of the right length

for(b in f){
  eigenvalues[b+1] <- Re(eigen(A2(0,b))$values[1])
}
# I used n0 into the A2 function since it still needs a population state vector as an input. However, it isn't taken into account in the linearization since a=0. The vector could really be anything, it just needs to still be an input.
```


Plot it up:
```{r}
length(eigenvalues) # check the length to be sure
plot(f,eigenvalues,type="p", xlab="f",ylab="dominant eigenvalue") # plot eigenvalues against f
abline(h=1,lty=2,col="red") # add a dashed red line at eigenvalue of 1
```


Time to look at the stable population size given different f values.

First, let's create an array of matrices (one for each f value):
```{r}
a=0.01
f <- 0:75
tf <- 500
nstg <- ncol(TS) # number of stages in the matrix
n0 <- c(2,rep(0,nstg-1)) # initial population vector
Ntf <- array(NA,c(length(f),nstg,tf)) # create an array of length(f) matrices
for(a in 1:length(f)) Ntf[a,,1] <- n0 # populate the array of matrices with the starting vector
```


Now, we run two for-loops to iterate across f values and through time:
```{r}
for(t in 1:(tf-1)){ # iterate across time
  for(b in f){ # iterate across f values
    Ntf[b+1,,t+1] <- A2(Ntf[b+1,,t],f=b)%*%Ntf[b+1,,t] # advance one time step, put it in the array entry matching to the value of f we're using, plus 1. This is because the first array entry is f=0, the second is f=1, etc.
  }
}
```


Now, plot it out:
```{r}
totals <- rep(NA,length(Ntf[,1,1])) # create a vector of the equilibrium pop size for each f value
for(i in 1:length(Ntf[,1,1])){ # calculate the total pop at the 500th time step for each f value
  totals[i] <- sum(Ntf[i,,500])
}

plot(1:length(totals),totals,type="p", xlab="f",ylab="stable pop size") # plot totals out


```

