---
title: "Assignment 7"
author: "Michael Culshaw-Maurer"
date: "2/28/2017"
output: html_document
---
```{r}
require(MASS)
```


Here's a modification of the lottery function we were given that also creates a kids matrix that will be used to calculate the stochastic growth rate.
```{r}
lottery.2=function(k=10,N0=rep(0.1,10),Sigma=diag(10),mu=seq(1,1.25,length=10),Tf=100,d=0.1){
  N=matrix(0,Tf,k) # matrix which will be filled in
  N[1,]=N0 # initial values
  K=matrix(NA,Tf-1,k) # kids matrix is 1 time step less than N matrix
  for(t in 2:Tf){ # starting the loop
    K[t-1,]=exp(mvrnorm(n=1,mu=mu,Sigma=Sigma)) # random draw creating the lognormal number of kids
    N[t,]=(1-d)*N[t-1,]+d*K[t-1,]*N[t-1,]/sum(K[t-1,]*N[t-1,]) # determining the next community state
  }
  return(list(N=N,K=K)) # returning both matrices 
}
```


Now, a function that runs through k species, putting them through the lottery function, and then calculating and returning the stochastic growth rates for each of the k species.
```{r}
calc.stoch <- function(k,Sigma,Tf,d){
  stoch <- numeric(length=k) # set up the vector for stochastic growth rates
  for(z in 1:k){ # start the loop through k species
    n0 <- rep((1/(k-1)),k) # initialize the starting values, except the one we set to 0
    n0[z] <- 0 # set the z entry to zero
    out <- lottery.2(N0 = n0,Sigma=Sigma,Tf=Tf,d=d) # run the lottery function
    K <- out$K # rename the variables for ease of use in formula below
    N <- out$N
    N <- N[-Tf,] # remove final N matrix time step so it matches the K matrix in size
    stoch[z] <- mean(log(1-d+d*K[,z]/rowSums(K*N))) # calculate each stochastic growth rate
  }
  return(stoch) # return the vector of stochastic growth rates
}
```


Now run the function to calculate stochastic growth rate for both of the Sigmas given in the assignment.
```{r}
dat1 <- calc.stoch(k=10,Sigma=diag(10),Tf=10000,d=0.1)
dat2 <- calc.stoch(k=10,Sigma=diag(10)*0.5+0.5,Tf=10000,d=0.1)
```


Combine our two simulations and make a barplot.
```{r}
dat <- rbind(dat1,dat2)
barplot(dat,beside = T,legend=c("diag(10)", "diag(10)*0.5+0.5"),args.legend = list(x="topleft"))
```

